How environmental drivers of spatial synchrony interact

### 2021-07-29 Paper/Paper.Rnw
started writing theory section of the main text

%new commands and so on
\providecommand{\keywords}[1]
{
  \small	
  \textbf{\textit{Keywords---}} #1
}
\newcommand{\mcP}{\mathcal{P}}
\newcommand{\mcB}{\mathcal{P}}

@@

%***Tentative author order
\author[a,b]{Daniel C. Reuman}
\author[c]{Max Castorani}
\author[d,e]{Tom Bell}
\author[f]{Kyle Cavanaugh}
\author[c]{Jonathan Walter}
\author[g]{Lawrence W. Sheppard}

\affil[a]{Department of Ecology and Evolutionary Biology and Kansas Biological Survey, University of Kansas}
\affil[b]{Laboratory of Populations, Rockefeller University}
\affil[c]{University of Virginia}
\affil[d]{WHOI}
\affil[e]{UCSB}
\affil[f]{UCLA}
\affil[g]{MBA}

@@

@@

@@

>>>>

%new commands and so on
\providecommand{\keywords}[1]
{
  \small	
  \textbf{\textit{Keywords---}} #1
}
\newcommand{\mcP}{\mathcal{P}}
\newcommand{\mcB}{\mathcal{P}}
\newcommand{\var}{\text{var}}
\newcommand{\cov}{\text{cov}}

@@

%Alternative titles
%Interactions among Moran effects
%Interacting Moran effects
%Interactions between environmental drivers of spatial synchrony

%***Tentative author order, who is last yet to be determined
\author[a,b]{Daniel C. Reuman}
\author[c]{Max Castorani}
\author[d,e]{Tom Bell}
\author[f]{Kyle Cavanaugh}
\author[c]{Jonathan Walter}
\author[a,g]{Lawrence W. Sheppard}

\affil[a]{Department of Ecology and Evolutionary Biology and Kansas Biological Survey, University of Kansas}
\affil[b]{Laboratory of Populations, Rockefeller University}
\affil[c]{Department of Environmental Sciences, University of Virginia}
\affil[d]{Applied Ocean Physics and Engineering Department, Woods Hole Oceanographic Institution}
\affil[e]{Earth Research Institute, University of California, Santa Barbara}
\affil[f]{Department of Geography, University of California, Los Angeles}
\affil[g]{Marine Biological Association of the United Kingdom}

@@

%Tentative paragraph outline written on my phone on 3/20/2021
%1) Importance of synchrony
%2) Classic problem of identifying causes of synchrony, but recently largely resolved due to new methods
%  -elaborate briefly on a couple methods
%3) in fact, it has recently been discovered that environmental causes of synchrony can interact, so that ...
%  -related to, but distinct from, interaction between environmental and dispersal causes of synchrony discussed, for instance by Kendal et al
%  -make sure you describe the two cases, cite the earlier papers, mention a few words about the two systems
%4) the simplest form of interacting moran effects can be demonstrated when a simple model, though elaboration of the true nature of this mechanism %will take the rest of this manuscript
%  -use Moran model
%  -but with two drivers
%  -moran theorem says correlation of pops equals correlation of sum of drivers, so work it out
%5) general idea of two noises lagged or time delayed relative to each other and their effects oppositely lagged - this may be where a pedagogical %fig comes in, or maybe it is just worsd here and then the pedagog fig comes in later, at end of theory section, in form of examples.
%6) paragraph on why we would expect interacting moran effects to be common, return to this point in the discussion
%7) final para on goals, manuscript structure, etc.
%  -G1) Elaborate a mathematical/statistical theory of interacting Moran effects which both allows both a detailed understanding 
%      of how the mechanism works and fecilitates statistical investigation of whether the effect is present in empirical systems.
%  -G2) Apply the theory and associated stats to revisiting kelp, the intention being to further elaborate the mechanism itself
%  -G3) Apply the theory and associated stats to plankton, the intention being to further elaborate the mechanism itself
%
%Notes:
%1) You need to make it clear the earlier papers already identified interacting Moran effects, so what are we doing?
%We are exploring theoreitically what the mechanism is, here, and how general it might be, and what different forms it 
%might take, using the theory, and also using the theory to further illuminate the two examples. The emphasis here
%is on illuminating the general mechanism of interacting Moran effects through the theory and its application, whereas
%the earlier papers focussed on the specific cases. The Fourier theory here serves to illuminate the general mechanism,
%whereas the earlier papers used wavelet statistics applied directly to data and did not really explore the general 
%mechanism. Need to work some of these thoughts in without seeming defensive.

@@

%Tentative broad outline:
%DONE 1) Intro spectral quantities, the spectrum, the cospectrum, and the cross spectrum of two time series/processes. And the spectral
%matrix for putting them all in one structure.
%2) Define your measures of timescale/frequency-specific synchrony and crosss-variable synchrony, explain why these make sense as
%ways to measure those things.
%3) Present the model and the theory
%4) Present "toy" examples and fig 1 (need to think a lot about these)
%
%Notes:
%Compared to the deer paper, this is a much deeper presentation of the theory in the main text. I'm not sure that's avoidable.
%For one thing, the theory is explicitly applied to get the results, later, and that was not the case for the deer paper.

\noindent Our main theoretical work requires a conceptual understanding of the \emph{spectrum} of an environmental 
or population time series, as well as of the \emph{cospectrum} and \emph{cross spectrum} of two time series.
If $w_i(t)$ is a stochastic process or time series measured at location $i$, the spectrum $S_{w_i w_i}(f)$ is a function
of frequency, $f$. For a periodic ocillation, $f$ can be defined as one over the timescale, or period, of the oscillation. 
The spectrum, $S_{w_i w_i}(f)$ is larger for frequencies at which $w_i(t)$ oscillates more. So, for example, a 
population $w_i(t)$ that shows strong oscillatory dynamics at $5$-year period will have a large value of 
$S_{w_i w_i}(f)$ for $f=1/5 \text{y}^{-1}$. The spectrum separates the overall variance of a time series into
contributions which occur on different timescales, in the sense that an integral of $S_{w_i w_i}(f)$ across all 
frequencies equals $\var(w_i)$. In a similar way, the cospectrum of two time series, $w_i(t)$ and $w_j(t)$,
is a function of $f$ that takes large values at frequencies for which oscillations in $w_i(t)$ and 
$w_j(t)$ are strong, and are strongly correlated, i.e., they have the same or similar phase. 
(Here, $j$ refers to another location where measurements were performed.) The cospectrum is 
the real part of the cross spectrum, $S_{w_i w_j}(f)$, which is a complex-valued function of frequency.
The cross-spectrum takes large-magnitude values at frequencies, $f$, for which the oscillatory components
of $w_i$ and $w_j$ are strong; and the complex phase of the cross spectrum at $f$ then quantifies
the phase lag between those oscillations. The notation $S_{ww}$ refers to the \emph{spectral matrix}, a 
matrix for which the $ij^{\text{th}}$ entry is $S_{w_i w_j}$. 
Spectral methods are standard tools in many fields of science, 
including ecology \citep{Vasseur2007,Reuman2008,Defriez2017}, and many background references are available 
[e.g., \cite{Brillinger2001}].

@@

\bibliographystyle{ecology_letters2}
\bibliography{refs}
